---
title: "PML Project Report"
author: "Krystle Sawyer"
date: "2023-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The goal of this project is to predict the manner in which 6 participants did their exercise (sitting, sitting down, standing, standing up, walking). This is the "classe" variable in the training set. Other variables may be utilized to predict with. This report describes how the model was built, how cross validation was used, what I think the expected out of sample error is, and why I made the choices I did. This prediction model will also be used to predict 20 different test cases. 


## Data 
The training data for this project is available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data is available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project comes from this source: http://groupware.les.inf.puc-rio.br/har. 



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
